# Results

```{r}
#| label: setup
#| include: false

library(tidyverse)
library(tidymodels)
library(gt)

pressures <- read_csv("_data/pressures.csv")

levels_habitat <- c("lake", "sea", "blank")
levels_treatment <- c("paper", "plastic", "control", "blank")

labels_habitat <-
  c(
    lake     = "Lake",
    sea      = "Sea",
    blank    = "Blank"
  )
labels_treatment <- 
  c(
    paper    = "Paper",
    plastic  = "Plastic",
    control  = "Control",
    blank    = "Blank"
  )

pressures <- 
  pressures |> 
  mutate(
    habitat = habitat |> fct(levels = levels_habitat),
    treatment = treatment |> fct(levels = levels_treatment),
    time = time / (60 * 24)
  )

blanks <- 
  pressures |> 
  filter(habitat == "blank") |> 
  filter(reactor != "B27") |> 
  summarize(
    blk_pressure = mean(pressure),
    .by = time
  )

pressures <- 
  pressures |> 
  left_join(blanks, by = join_by(time)) |> 
  mutate(adj_pressure = pressure - blk_pressure)

tm <- 25 + 273.15 # sample temperature in Kelvin
t0 <- 273.15 # reference temperature in Kelvin
a <- 0.03103 # Bunsen absorption coefficient
r <- 83.144 # gas constant in L.hpa/mol.K
mo2 <- 32000 # oxygen gas molecular weight in mg/mol
vt <- 995 # reactor volume in mL
sv <- 260 # sample volume in mL.

pressures <- 
  pressures |> 
  mutate(bod = mo2 / (r * tm) * ((vt - sv) / sv + a * tm / t0) * -adj_pressure)

pressures <- 
  pressures |> 
  mutate(outliers = if_else(reactor %in% c("L13", "L17", "L18"), TRUE, FALSE))
```

## Blank evolution

Differences in pressure are adjusted offsetting from the mean value in the empty reactors---also called blanks. During the first couple of days, all blanks rose to approximately 9 hPa over the starting pressure (@fig-blanks). The response is likely due to warming air inside the reactor, because the incubator chamber was at 25°C, which was slightly above room temperature.

Reactor B27 lost pressure early in the experiment and dropped to zero, this behavior implies evident sealing defects and the reactor was not considered for the analysis.

Our goal is to measure differences in pressure due to oxygen consumption ---or CO-2- production---, so we neutralized the effect of temperature by adjusting the pressure difference with the formula:

$p_{OD} = p - \overline{p}_{blanks}$

Were $p_{OD}$ is the pressure difference driven by oxygen consumption, $p$ is the relative pressure of the reactors containing benthic habitats, and $\overline{p}_{blanks}$ is the average relative pressure measured in empty reactors ---excluding reactor B27.

:::{#fig-blanks}
```{r}
pressures |> 
  filter(time %% 1 == 0) |> 
  filter(treatment == "blank") |> 
  ggplot(aes(x = time, y = pressure, color = reactor)) +
  geom_point() +
  geom_line() +
  labs(x = "Time (days)", y = "Relative pressure (hPa)", color = "Reactor") +
  theme_classic()
```
Relative pressure evolution inside empty reactors. Reactor B27 dropping to zero implies that it was not properly sealed.
:::

::: {#fig-habitat-characterization-reactors}
```{r}
pressures |> 
  filter(time %% 1 == 0) |> 
  filter(treatment == "control") |> 
  ggplot(aes(x = time, y = bod, color = habitat, group = reactor)) +
  geom_line(aes(linetype = habitat)) +
  geom_point(aes(shape = habitat), size = 2) +
  scale_linetype_manual(labels = labels_habitat, values = c("dashed", "solid")) +
  scale_color_discrete(labels = labels_habitat) +
  scale_shape_manual(labels = labels_habitat, values = c(1, 17)) +
  labs(x = "Time (days)", y = "Oxygen consumption (mg/L)") +
  theme_classic() +
  theme(
    legend.position = "inside",
    legend.position.inside = c(.1, .9),
    legend.background = element_rect(color = "black"),
    legend.title = element_blank()
  )
```
Oxygen evolution inside untreated reactors.
:::

::: {#fig-treatments}
```{r}
pressures |>
  filter(time %% 5 == 0) |> 
  filter(treatment != "blank") |> 
  ggplot(aes(x = time, y = bod, shape = treatment, color = treatment)) +
  geom_line(
    aes(linetype = treatment),
    stat = "summary", fun = mean,
    position = position_dodge(width = 3),
    show.legend = F
  ) +
  geom_pointrange(
    stat = "summary", fun.data = mean_se,
    position = position_dodge(width = 3)
  ) +
  facet_wrap(~ habitat, scales = "free_y", labeller = as_labeller(labels_habitat)) +
  scale_shape_manual(labels = labels_treatment, values = c(0, 16, 1)) +
  scale_color_discrete(labels = labels_treatment) +
  scale_linetype_manual(values = c("dotted", "solid", "dashed")) +
  labs(x = "Time (days)", y = "Oxygen consumption (mg/L)") +
  theme_classic() +
  theme(
    legend.position = "inside",
    legend.position.inside = c(.12, .8),
    legend.background = element_rect(color = "black"),
    legend.title = element_blank()
  )
```
Summary evolution of biological oxygen demand for each treatment. Points represent reactor averages and bars standard error range.
:::

```{r}
pressures |>
  filter(time %% 1 == 0) |> 
  filter(treatment != "blank") |> 
  ggplot(aes(x = time, y = bod, color = treatment, linetype = treatment)) +
  geom_line(aes(group = reactor)) +
  facet_wrap(~ habitat, scales = "free_y", labeller = as_labeller(labels_habitat)) +
  scale_linetype_manual(labels = labels_treatment, values = c("dashed", "solid", "dotted"), name = NULL) +
  scale_color_discrete(labels = labels_treatment, name = NULL) +
  labs(x = "Time (days)", y = "Oxygen consumption (mg/L)") +
  theme_classic() +
  theme(legend.position = "inside", legend.background = element_rect(color = "black"), legend.position.inside = c(.1, .8))
```

```{r}
pressures |> 
  filter(time %% 1 == 0) |> 
  filter(treatment == "control") |> 
  ggplot(aes(x = time, y = bod)) +
  geom_pointrange(
    aes(shape = habitat),
    stat = "summary",
    position = position_dodge(width = 1),
    fun.data = mean_se
  ) +
  scale_shape_manual(values = c(1, 16)) +
  scale_linetype_manual(values = c("dotted", "solid")) +
  labs(x = "Time (days)", y = "Oxygen consumption (mg/L)") +
  theme_classic() +
  theme(
    legend.position = "inside",
    legend.position.inside = c(.1, .9),
    legend.box.background = element_rect(color = "black"),
    legend.title = element_blank()
  )
```


